---
title: "Statistical assignment 4"
author: "Benjamin Machon 680011539"
date: "02/03/2020"
output: github_document
---

```{r setup, include=FALSE}
# Please note these options.
# This tells R Markdown that we want to show code in the output document.
knitr::opts_chunk$set(echo = TRUE)
# Switching off messages in the output document. 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
# Switching on caching to make things faster (don't commit cache files on Github).
knitr::opts_chunk$set(cache = TRUE)
```

In this assignment you will need to reproduce 5 ggplot graphs. I supply graphs as images; you need to write the ggplot2 code to reproduce them and knit and submit a Markdown document with the reproduced graphs (as well as your .Rmd file).

First we will need to open and recode the data. I supply the code for this; you only need to change the file paths.

    ```{r}
library(tidyverse)
Data8 <- read_tsv("/Users/benmachon/Downloads/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")
Data8 <- Data8 %>%
        select(pidp, h_age_dv, h_payn_dv, h_gor_dv)
Stable <- read_tsv("/Users/benmachon/Downloads/UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab")
Stable <- Stable %>%
        select(pidp, sex_dv, ukborn, plbornc)
Data <- Data8 %>% left_join(Stable, "pidp")
rm(Data8, Stable)
Data <- Data %>%
        mutate(sex_dv = ifelse(sex_dv == 1, "male",
                           ifelse(sex_dv == 2, "female", NA))) %>%
        mutate(h_payn_dv = ifelse(h_payn_dv < 0, NA, h_payn_dv)) %>%
        mutate(h_gor_dv = recode(h_gor_dv,
                         `-9` = NA_character_,
                         `1` = "North East",
                         `2` = "North West",
                         `3` = "Yorkshire",
                         `4` = "East Midlands",
                         `5` = "West Midlands",
                         `6` = "East of England",
                         `7` = "London",
                         `8` = "South East",
                         `9` = "South West",
                         `10` = "Wales",
                         `11` = "Scotland",
                         `12` = "Northern Ireland")) %>%
        mutate(placeBorn = case_when(
                ukborn  == -9 ~ NA_character_,
                ukborn < 5 ~ "UK",
                plbornc == 5 ~ "Ireland",
                plbornc == 18 ~ "India",
                plbornc == 19 ~ "Pakistan",
                plbornc == 20 ~ "Bangladesh",
                plbornc == 10 ~ "Poland",
                plbornc == 27 ~ "Jamaica",
                plbornc == 24 ~ "Nigeria",
                TRUE ~ "other")
        )
    ```

Reproduce the following graphs as close as you can. For each graph, write two sentences (not more!) describing its main message.  

1. Univariate distribution (20 points).

    ```{r}
    ggplot(data = Data, aes(x = h_payn_dv)) +
        geom_freqpoly() + 
        labs(y = "Number of respondents", x = "Net Monthly Pay")
    
    ```
This plot shows that just over £1000 is the most common monthly pay for the respondents in the survey. After a small spike at around 5,500 the number of respondents remains at 0.


2. Line chart (20 points). The lines show the non-parametric association between age and monthly earnings for men and women.

    ```{r}
# Age for both sexes
byAge <- Data %>%
  group_by(h_age_dv) %>%
  summarise(
    monthly = mean(h_payn_dv, na.rm = TRUE)
  )
byAge %>%
        ggplot(aes(x = h_age_dv, y = monthly)) +
        geom_smooth() +
        xlim(10,60) +
        xlab("Age") +
        ylab("Monthly Earnings")

# splitting up sex

byAgeSex <- Data %>%
  filter(!is.na(sex_dv)) %>%
  group_by(h_age_dv, sex_dv) %>%
  summarise(
monthly = mean(h_payn_dv, na.rm = TRUE)
  )

byAgeSex %>%
    ggplot(aes(x = h_age_dv, y = monthly, colour = sex_dv)) +
    geom_smooth() +
    xlim(15,65) +
    ylim(0,2500) +
    labs(x = "Age", y = "Monthly earnings", colour = "Sex")

# I struggled to change the colour and pattern of the lines.

    ```
This line graph shows us the relationship between the age of a person and their monthly earnings. Both me and women's average monthly wages increase until the age of 45 before they plateau and decrease slightly. However, it remains noticeable that a man's avrage monthly earnings remain significantly higher than that of a woman throughout their lifetime

3. Faceted bar chart (20 points).

    ```{r}
    byNat <- Data %>%
  group_by(placeBorn, sex_dv) %>%
  summarise(
    monthly = median(h_payn_dv, na.rm = TRUE)
  ) %>%
        filter(!is.na(sex_dv)) %>%
    filter(!is.na(placeBorn)) %>%
    
  ggplot(aes(x = sex_dv, y = monthly)) +
  geom_bar(stat = "identity") +
        labs(x = "Sex", y = "Median monthly net pay") +
facet_wrap(~ placeBorn)
    
    byNat
    ```
This graph helps to show the distribution of pay between men and women, grouped by the country in which they were born. In every case, men earn more than women on (median) average per month


4. Heat map (20 points).

    ```{r}
   HeatRegion <- Data %>%
        select(h_gor_dv, placeBorn, h_age_dv) %>%
        filter(!is.na(h_gor_dv)) %>%
        filter(!is.na(placeBorn)) %>%
        
        
    ggplot(aes(x = h_gor_dv, y = placeBorn, fill = h_age_dv)) +
    geom_tile() +
        labs(x = "Region", y = "Country of birth", fill = "Mean age") +
        theme(axis.text.x = element_text(angle = 90)) 
        
    HeatRegion
    
    # I attempted to use the lims() function to alter the scale of the Mean Age but was unsuccessful
    ```
This graph is supposed to show the average age of respondents according to where they were born and where they now reside within the UK. However, due to what I believe as problems with the scale means that the values are inaccurate compared to the heat map on the question. Therefore it is difficult to analyse.

5. Population pyramid (20 points).

    ```{r}
    
    PP <- Data %>%
        select(h_age_dv, sex_dv) %>%
        filter(!is.na(sex_dv)) %>%
        
        ggplot(aes(x = h_age_dv, fill = sex_dv)) +
        geom_bar(subset = "identity") +
        coord_flip() +
        labs(x = "Age", y = "n", fill = "Sex") +
        ylim(-400, 500)
    
    PP
    
    ```
    I am struggling to reflect the graph onto the opposite axis.
    
    
    